Conditional selectivity for triplestores

### 2012-09-19 conditional/trunk/selectivity-basic.tex
minor updates
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@385 43ef1190-90fd-4dd7-a769-d0121f02b80e

From elementary statistics, we know that the intersection probability
can be written as
\begin{eqnarray}
Pr(A_1 \cap A_2  \cap \ldots \cap A_n) &=& 
  Pr(A_n | A_1 \cap A_2  \cap \ldots \cap A_{n-1}) \nonumber\\
  &&\cdot Pr(A_{n-1} | A_1 \cap A_2  \cap \ldots \cap A_{n-2}) \cdot \ldots \nonumber\\
  && \cdot Pr(A_2 | A_1) \cdot Pr(A_1) ,
\end{eqnarray}
where the $|$ symbol is read as ``given'', e.g. ``given a predicate
\rdfterm{rdf:type} what is the probability that the object is \rdfterm{foaf:Person}?''

@@

Uniform distributions have no impact on join order\todo{argue}.
\todo{Use parameterized stats for datatype properties?}

@@

Allthough a careful estimate of the selectivity in situations where we
have e.g. $\sel{O = ? | P = p \cap S = ?}$ would require a two
dimensional distribution function, the most significant impacts on
join order is when the distribution is highly skewed\todo{argue}. A
hint to the optimizer in such cases could be achieved by the addition
of properties like \rdfterm{void:mostFrequentObjects},
\rdfterm{void:mostFrequentSubjects},
\rdfterm{void:leastFrequentObjects} and 
\rdfterm{void:leastFrequentSubjects} to property
partitions.\todo{mention the dc:language case in Sublima?}

>>>>

From elementary statistics, we know that the intersection probability
can be written as
\begin{eqnarray}
Pr(A_1 \cap A_2  \cap \ldots \cap A_n) &=& 
  Pr(A_n | A_1 \cap A_2  \cap \ldots \cap A_{n-1}) \nonumber\\
  &&\cdot Pr(A_{n-1} | A_1 \cap A_2  \cap \ldots \cap A_{n-2}) \cdot \ldots \nonumber\\
  && \cdot Pr(A_2 | A_1) \cdot Pr(A_1) ,
\end{eqnarray}
where the $|$ symbol is read as ``given'', e.g. ``given a predicate
\rdfterm{rdf:type} what is the probability that the object is
\rdfterm{foaf:Person}?'' This is known as the ``chain rule''.

@@

Uniform distributions have little interesting impact on join
order\todo{argue}. \todo{fallback estimator: distinctSubjects/triples, etc?}
\todo{Use parameterized stats for datatype properties?}

@@

Allthough a careful estimate of the selectivity in situations where we
have e.g. $\sel{O = ? | P = p \cap S = ?}$ would require a two
dimensional joint distribution function, the most significant impacts on
join order is when the distribution is highly skewed\todo{argue}. A
hint to the optimizer in such cases could be achieved by the addition
of properties like \rdfterm{void:mostFrequentObjects},
\rdfterm{void:mostFrequentSubjects},
\rdfterm{void:leastFrequentObjects} and 
\rdfterm{void:leastFrequentSubjects} to property
partitions.\todo{mention the dc:language case in Sublima?}
